---
title: "U.S Incarceration Systems Investigation"
author: "Hiba Abbas"
date: '2022-11-01'
output: html_document
---

```{r echo=FALSE, eval=TRUE, message=FALSE}
source("~/Documents/info201/assignments/a4-HibbaAbas/source/a4-helpers.R")
source("~/Documents/info201/assignments/a4-HibbaAbas/source/analysis.R")

```

## Abstract 
This report uses r and Ventura data to investigate the U.S jail systems.The Ventura data used has recorded data from 1970 to 2018. In this report the prisons' care for the well being of its inmates are examined through the recorded death counts of inmates, the racial disparities in the jail populations, and more are included.

## Introduction 
America has the highest jailing rate in the world. Yada Yada Yada

## Data Summary 
America's prison systems are not good, the health of the incarcerated is neglected. From 2000 to 2013 there were **`r total_jail_deaths_pretty`** deaths across all prisons in the nation. The amount of deaths would increase with each year, in 2000 there were approximatly **`r jail_deaths_2000_pretty`** and in 2013 there were **`r jail_deaths_2013_pretty`**, meaning there was an **`r prettyNum(jail_deaths_2013 - jail_deaths_2000, big.mark=",", scientific=FALSE)`** increase in deaths over thirteen years. 


## Section 3

```{r echo=FALSE}
plot_jail_pop_for_us()
```

## Section 4

```{r echo=FALSE}
states <- c("WA", "CA", "TX")
plot_jail_pop_by_states(states)
```

## Section 5
```{r echo=FALSE}
plot_jail_pop_by_race()
```

_All the very best with your work!_

